Supplemental material: Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-02-23 Manuscript/HI_Model_ms.Rmd
updated manuscript files

### *4.1 Impact of different climate and land-use pathways* 
By comparing model results of Hawaiian ecosystem carbon balance under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and disturbances such as wildland fire (figure 5). Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest cumulative impact, resulting in a median of 26 additional Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period. Long-term reductions in the intensity of land-use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions. Switching from the ‘high’ to the ‘low’ land‐use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100. The combination of global climate mitigation and a reduction in land‐use conversions had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C). Notably, the relative impact of reducing emissions from land-use change was much greater under the high radiative forcing pathway (RCP 8.5). Cumulative NBP increased by 130% when switching from the high to low land-use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land-use under RCP 4.5.

>>>>

### *4.1 Impact of different climate and land-use pathways* 
By comparing ecosystem carbon balance estimates under different scenario combinations, we were able to assess the relative impact of both global emissions reductions and regional actions to reduce emissions from land use, land use change, and disturbances such as wildland fire (figure 5). Global adherence to a lower emissions trajectory (i.e., switching from RCP 8.5 to RCP 4.5) had the largest cumulative impact, resulting in a median of 26 additional Tg C sequestered by Hawaiian ecosystems over the 90-year simulation period. Long-term reductions in the intensity of land-use change also consistently led to an increase in ecosystem carbon sequestration, but to a lesser degree than global emissions reductions. Switching from the ‘high’ to the ‘low’ land‐use scenario resulted in a median cumulative retention of an additional 11.6 Tg C in Hawaiian ecosystems by 2100. The combination of global climate mitigation and a reduction in land‐use conversions had the largest potential benefit to ecosystem carbon sequestration, reducing cumulative net losses by over 400% (37.7 Tg C). Notably, the relative impact of reducing emissions from land-use change was much greater under the high radiative forcing pathway (RCP 8.5). Cumulative NBP increased by 130% when switching from the high to low land-use scenario under RCP 8.5, as opposed to a 37% cumulative increase in NBP when switching from high to low land-use under RCP 4.5.

### 2021-02-23 Manuscript/HI_Model_supp.Rmd
updated manuscript files

This study used a new spatial database of wildland fire perimeters in the Hawaiian Islands for the years 1999-2019 using data compiled from the Monitoring Trends in Burn Severity (MTBS) program, the Hawaiʻi Wildfire Management Organization, the U.S. National Park Service, the Oʻahu Army Natural Resources Program, and independent remote sensing analyses using publicly available LandSat data. This new wildland fire spatial database was used to calculate annual area burned (Figure S2), wildland fire probabilities by Moisture Zone and State Class, and wildland fire size distributions.

>>>>

This study used a new spatial database of wildland fire perimeters on the main Hawaiian Islands from 1999-2019 to calculate annual area burned (Figure S2), wildland fire probabilities by Moisture Zone and State Class, and wildland fire size distributions. This new wildland fire spatial database compiles prior mapping efforts and data collections with new fire perimeter data mapped directly by Dr. Clay Trauernicht in the Department of Natural Resources and Environmental Management at the University of Hawaiʻi at Mānoa. The goal of this effort was to locate and map all fires greater than or equal to 20 hectares, but some smaller fires were included as detected in imagery. Fire perimeters for the years 2002-2011 were primarily from the U.S. Geological Survey's Monitoring Trends in Burn Severity (MTBS; https://www.mtbs.gov). Hawaii Wildfire Management Organization (www.hawaiiwildfire.org) provided ground-based, GPS-mapped fire perimeters primarily from the Kona and Kohala regions of Hawaiʻi Island. The U.S. National Park Service provided ground-based, GPS-mapped fires records from Hawaiʻi Volcanoes National Park. The Oʻahu Army Natural Resource Program provided ground-based, GPS-mapped fires on Oʻahu. All other fires were mapped directly by Dr. Trauernicht using LANDSAT and Sentinel-2 satellite imagery. For 2002-2011, USGS MTBS data were prioritized in the case of any duplicate record. Dr. Trauernicht mapped all fires from 2012 onwards in addition to filling missing records, where possible, using LANDSAT or Sentinel-2 imagery.
